Iterative Reduction: Parsing and Interpreting Code Without Trees

### 2024-12-16 Paper/paper.tex
comments for states

We define the following four base sets:
\begin{enumerate}
    \item $\values$ the \textit{universe of values}; these are all the internal values an object may have.
    \item $\termp$ the set of \textit{printable terms}; this is the set of all printable tokens a programmer may type.
    \item $\terms$ the set of \textit{type terms}; these are terms which correspond to types in the classical sense.
    \item $\terma$ the set of \textit{abstract terms}; these are intermediate terms (like $\OP\NUM$ from the previous section).
\end{enumerate}
The sets $\termp,\terms,\terma$ are disjoint, and we place no such restriction on $\values$.

@@

@@

So if $\Sigma=\set{\NUM_0,\LIST_1,\PRODUCT_\infty}$ (the subscripts are the arities of each base type) and $\mathcal A=\set{\OP}$, then
\[ \NUM,\LIST\NUM,\PRODUCT(\LIST\NUM\NUM)\in\terms,\qquad \OP\LIST\NUM\in\terma \]
For readability we can put parentheses and commas where necessary, but importantly these sets have the \textit{unique reconstruction property}: it is possible to compute their construction from the
string itself (i.e. a string cannot be constructed from two different patterns).

Now we define the following sets:
\begin{enumerate}
    \item $\term=\termp\cup\terms\cup\terms$, the set of \textit{terms},
    \item $\termi=\terms\cup\terma$, the set of \textit{internal terms},
    \item $\expi=\termi\times\values$, the set of \textit{internal expressions},
    \item $\exp=\termi\cup\termp$, the set of \textit{expressions}.
\end{enumerate}
Finally we define $\oe=\exp\times\bbZ$.
Elements of $\oe$ will be written like $\sigma_n(v)$ as before; for printable terms we act as if the value is empty.

The \textit{initial priority function} is a function $\pi\colon\termp\longto\bbZ$.
This can be naturally extended to $\pi\colon\termp^*\longto\oe^*$.

Finally, our \textit{initial $\beta$-reducer} is a function ($\state$ is a set of functions $\termp\partial\expi$ which we define later)
\[ \hb\colon\termi\times\term^\epsilon \partial \termi^\epsilon\times(\bbZ\times\bbZ\to\bbZ)\times(\values\times\values\times\state\partial\values\times\termp^*\times\state) \]
and we extend this to a \textit{derived $\beta$-reducer}, a function
\[ \beta\colon\oe^*\times\state \partial \oe^*\times\state \]
inputs in $\oe^*\times\state$ will be written as $\gen\xi[\bar s]$ for $\xi\in\oe^*$ and $\bar s\in\state$.
We define $\beta^*\gen\xi[\bar s]$ to be the value obtained by composing $\beta$ with itself of $\gen\xi[\bar s]$ until the convergence of $\xi$.
Meaning
\begin{multline*}
    \beta^* = \beta^n\gen\xi[\bar s]\\
    \hbox{where $n=\min\set{n\geq1}[{\beta^n\gen\xi[\bar s]=\gen{\xi'}[\bar s'],\beta\gen{\xi'}[\bar s']=\gen{\xi'}[\bar s'']}]$}
\end{multline*}
We define $\beta\gen\xi[\bar s]$ as follows, where $\xi=\sigma^1_{p_1}(v_1)\cdots\sigma^n_{p_n}(v_n)$.
\begin{enumerate}
    \item if $\sigma^1\in\termp$ then
    \[ \beta\gen\xi[\bar s] = \gen{s(\sigma^1)_{p_1}\sigma^2_{p_2}(v_2)\cdots\sigma^n_{p_n}(v_n)}[\bar s] \]
    \item if $\hb(\sigma^1,\epsilon)=(\tau,\rho,f)$ is defined, then compute $f(v_1,\_,s)=(u,\xi',s')$ and $\beta^*\gen{\pi\xi'}[\bar s']=\gen{\xi''}[\bar s'']$, and
    \[ \beta\gen\xi[\bar s] = \gen{\tau_{\rho(p_1,\_)}\xi''\sigma^2_{p_2}(v_2)\cdots\sigma^n_{p_n}(v_n)}[\bar s''] \]
    \item if $p_1\geq p_2$ and $\hb(\sigma^1,\sigma^2)=(\tau,\rho,f)$ is defined, then compute $f(v_1,v_2,s)=(u,\xi',s')$ and $\beta^*\gen{\pi\xi'}[\bar s']=\gen{\xi''}[\bar s'']$, and
    \[ \beta\gen\xi[\bar s] = \gen{\tau_{\rho(p_1,p_2)}\xi''\sigma^3_{p_3}(v_2)\cdots\sigma^n_{p_n}(v_n)}[\bar s''] \]
    \item otherwise compute $\beta\gen{\sigma^2_{p_2}(v_2)\cdots\sigma^n_{p_n}(v_n)}[\bar s]=\gen{\xi'}[\bar s']$ and
    \[ \beta\gen\xi[\bar s] = \gen{\sigma^1_{p_1}(v_1)\xi'}[\bar s'] \]
\end{enumerate}

>>>>

We define the following four base sets:
\begin{enumerate}
    \item $\values$ the \textit{universe of values}; these are all the internal values an object may have.
    \item $\termp$ the set of \textit{printable terms}; this is the set of all printable tokens a programmer may type.
    \item $\terms$ the set of \textit{type terms}; these are terms which correspond to types in the classical sense.
    \item $\terma$ the set of \textit{abstract terms}; these are intermediate terms (like $\OP\NUM$ from the previous section).
\end{enumerate}
The sets $\termp,\terms,\terma$ are disjoint, and we place no such restriction on $\values$.
\yoni{The comment about no restrictions for $\values$ is unclear.}

@@

\yoni{I don't think this is very standard. Perhaps you can define
 it inductively: Every element of $\Sigma^0$ is in $T_{\Sigma}$,
  and for eac $n$, if $t_i\in T_{\Sigma}$ for each $1\leq i \leq n$
  then $f t_1 \ldots t_n \in T_{\Sigma}$ for each $f\in\Sigma^{n}$,
  and then for each $g\in\Sigma^{\infty}$ and any $n$, similarly...
Do we actually need $\Sigma^{\infty}$ BTW? }

@@

\begin{example}
If $\Sigma^{0}=\{\NUM\}$,
$\Sigma^{1}=\{\LIST\}$, and $\Sigma^{\infty}=\{\PRODUCT\}$
and $\Sigma=\Sigma^0\cup\Sigma^1\cup\Sigma^\infty$,
 and $\mathcal A=\set{\OP}$, 
\yoni{plz specify the "arity" of $\OP$.}
 then
\[ \NUM,\LIST\NUM,\PRODUCT(\LIST\NUM\NUM)\in\terms,\qquad \OP\LIST\NUM\in\terma \]
\yoni{I think it is better to put a space between these, especially
in $\LIST\NUM\NUM$.}
\end{example}
For readability we can put parentheses and commas where necessary, but importantly these sets have the \textit{unique reconstruction property}:
\yoni{If I understand correctly you are assuming this property, right? If so, please say so.}
it is possible to compute their construction from the
string itself (i.e. a string cannot be constructed from two different patterns).

Now we define the following sets:
\begin{enumerate}
    \item $\term=\termp\cup\terms\cup\terms$, 
\yoni{Should the last $\terms$ be $\terma$?}
    the set of \textit{terms},
    \item $\termi=\terms\cup\terma$, the set of \textit{internal terms},
    \item $\expi=\termi\times\values$, the set of \textit{internal expressions},
    \item $\exp=\termi\cup\termp$, the set of \textit{expressions}.
\end{enumerate}
Finally we define $\oe=\exp\times\bbZ$.
Elements of $\oe$ will be written like $\sigma_n(v)$ as before; 
\yoni{But elements of $\oe$ seems like pairs to me, not triplets.}
for printable terms we act as if the value is empty.
\yoni{Please clarify what does acting as if... means.}

The \textit{initial priority function} is a function $\pi\colon\termp\longto\bbZ$.
This can be naturally extended to $\pi\colon\termp^*\longto\oe^*$.
\yoni{please show how.}

Finally, our \textit{initial $\beta$-reducer} is a function ($\state$ is a set of functions $\termp\partial\expi$ which we define later)
\yoni{Using something and defining it later should be avoided...}

\yoni{Generally, this section is hard to follow. I suggest adding plenty of examples, after every new definition / concept, please add an appropriate small example.}

\[ \hb\colon\termi\times\term^\epsilon \partial \termi^\epsilon\times(\bbZ\times\bbZ\to\bbZ)\times(\values\times\values\times\state\partial\values\times\termp^*\times\state) \]
and we extend this to a \textit{derived $\beta$-reducer}, a function
\[ \beta\colon\oe^*\times\state \partial \oe^*\times\state \]
inputs in $\oe^*\times\state$ will be written as $\gen\xi[\bar s]$ for $\xi\in\oe^*$ and $\bar s\in\state$.
We define $\beta^*\gen\xi[\bar s]$ to be the value obtained by composing $\beta$ with itself of $\gen\xi[\bar s]$ until the convergence of $\xi$.
Meaning
\begin{multline*}
    \beta^* = \beta^n\gen\xi[\bar s]\\
    \hbox{where $n=\min\set{n\geq1}[{\beta^n\gen\xi[\bar s]=\gen{\xi'}[\bar s'],\beta\gen{\xi'}[\bar s']=\gen{\xi'}[\bar s'']}]$}
\end{multline*}
We define $\beta\gen\xi[\bar s]$ as follows, where $\xi=\sigma^1_{p_1}(v_1)\cdots\sigma^n_{p_n}(v_n)$.
\begin{enumerate}
    \item if $\sigma^1\in\termp$ then
    \[ \beta\gen\xi[\bar s] = \gen{s(\sigma^1)_{p_1}\sigma^2_{p_2}(v_2)\cdots\sigma^n_{p_n}(v_n)}[\bar s] \]
    \item if $\hb(\sigma^1,\epsilon)=(\tau,\rho,f)$ is defined, then compute $f(v_1,\_,s)=(u,\xi',s')$ and $\beta^*\gen{\pi\xi'}[\bar s']=\gen{\xi''}[\bar s'']$, and
    \[ \beta\gen\xi[\bar s] = \gen{\tau_{\rho(p_1,\_)}\xi''\sigma^2_{p_2}(v_2)\cdots\sigma^n_{p_n}(v_n)}[\bar s''] \]
    \item if $p_1\geq p_2$ and $\hb(\sigma^1,\sigma^2)=(\tau,\rho,f)$ is defined, then compute $f(v_1,v_2,s)=(u,\xi',s')$ and $\beta^*\gen{\pi\xi'}[\bar s']=\gen{\xi''}[\bar s'']$, and
    \[ \beta\gen\xi[\bar s] = \gen{\tau_{\rho(p_1,p_2)}\xi''\sigma^3_{p_3}(v_2)\cdots\sigma^n_{p_n}(v_n)}[\bar s''] \]
    \item otherwise compute $\beta\gen{\sigma^2_{p_2}(v_2)\cdots\sigma^n_{p_n}(v_n)}[\bar s]=\gen{\xi'}[\bar s']$ and
    \[ \beta\gen\xi[\bar s] = \gen{\sigma^1_{p_1}(v_1)\xi'}[\bar s'] \]
\end{enumerate}
